---
title: "Seismic Reflection Data"
author: "Ibrahim Ilhan"
date: "8/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Vibhor's comment

```{r}

print("Hello Owner")
```

# add
# Packages that will be utilized

```{r}
suppressPackageStartupMessages({
  library(leaflet)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
})
```

# Load salmon data from the KNB

```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/knb.92020.1"
# data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"
esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)

head(esc)
```

# Prapare the dataset for calculating medians

```{r}
median_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  group_by(Species, SASAP.Region, Year, Location) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
  group_by(Species) %>% 
  summarize(median_escapement = median(escapement))

head(median_esc)
```
# Plot the escapement as a bar chart

```{r}
ggplot(median_esc, aes(Species, median_escapement)) +
  # ggplot(file, median_esc; x, median_esc; y, Species) | aes, Aesthetics
  geom_col() +
  # bar charts # help, ?geom_
  coord_flip()
  # rotate the coordinates
```

# Now work on building a map
We are starting from daily data escapement data again, and finding the uique locations at which escapement was measured

# Find unique location point
Make sure to remember to deal with missing values in the location field

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
# drop all that doesn't have lat. long.
  datatable(locations)
```

# Draw the map
Use leaflet to make interactive map.

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  # addTiles backgroung 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
  # 
```

```{r}
locs <- locations %>% mutate(Longitude = abs(Longitude) * -1)

leaflet(locs) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
```







