---
title: "tidying Data in R"
author: "Ibrahim Ilhan"
date: "8/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Read in libraries

Calling a masked function

package::functionName
example:: filter()


```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
```
Read in data from [Mike Byerly. 2016. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/#view/df35b.304.2)

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```


```{r}
catch_df %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
  head()
```

```{r}
catch_df %>% 
  select(-All, -notesRegCode) %>% 
  head()
```

```{r}
catch_df <- catch_df %>% 
  select(-All, -notesRegCode)
  # select the other columns, not All and notesRegCode
head(catch_df)
```
Move from wide to long format (using gather), and rename the cacth column to reflect units

```{r}
catch_df <- catch_df %>% 
  gather(key = "species", value = "catch", -Region, -Year) %>% 
  rename(catch_thousands = catch)

  head(catch_df)
```
```{r}
catch_integers <- as.integer(catch_df$catch_thousands)

i <- which(is.na(catch_integers) == T)

catch_df$catch_thousands[i]
```
[1] "I"

1) turn "I" into 1 in the catch_thousands column
2) coerce catch_thousands into integer type
3) convert catch_thousands to catch

```{r}
catch_df <- catch_df %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000)

tail(catch_df)

```

# Split-Apply-Combine

```{r}
catch_summarized <- catch_df %>%
  group_by(Region) %>%
  summarise(mean_catch = mean(catch))

head(catch_summarized)
```
```{r}
catch_mean <- catch_df %>%
  group_by(Region) %>%
  summarise(catch_mean = mean(catch), catch_stdev = sd(catch))

head(catch_mean)
```


```{r}
catch_mean <- catch_df %>%
  group_by(Region) %>%
  # calculate mean, sd, n
  summarise(catch_mean = mean(catch, na.rm = T), 
            catch_stdev = sd(catch, na.rm = T), 
            num_obs = n())

head(catch_mean)
```
```{r}
catch_mean <- catch_df %>%
  group_by(Region, species) %>%
  # calculate mean, sd, n
  summarise(catch_mean = mean(catch, na.rm = T), 
            catch_stdev = sd(catch, na.rm = T), 
            num_obs = n())

head(catch_mean)
```

```{r}
catch_mean <- catch_df %>%
  group_by(Region, species) %>%
  # calculate mean, sd, n
  summarise(catch_mean = mean(catch, na.rm = T), 
            catch_stdev = sd(catch, na.rm = T), 
            num_obs = n()) %>% 
  # fileter for Chinook
  filter(species == "Chinook") %>% 
  filter(catch_mean > 2000)

head(catch_mean)
```

```{r}
catch_year <- catch_df %>%
  filter(Year > 1990 & species == "Chinook") %>% 
  # filter(Year > 1990 | Year < 1900) %>% 
  group_by(Year, species) %>%
  # calculate mean, sd, n
  summarise(catch_sum = sum(catch))

print(catch_year)
```

Reading region_defs data from website

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)
head(region_defs)
```

```{r}
region_defs <- region_defs %>% 
  select(code, mgmtArea)

  head(region_defs)
```

Merging two different files (catch_df and region_defs) based on Region and code columns, which contain the same information.. So the merged file has Region, and the code is removed/replacced. 

```{r}
catch_joined <- left_join(catch_df, region_defs, by = c("Region" = "code"))

head(catch_joined)
```

